$\pi$SOD-M: A Methodology for Building Service-Oriented Applications in the Presence of Non-Functional Properties

### 2014-03-14 JournalToSubmit2013/modelingWithPISODM.tex
New version for the section transformation

 %Transformations in the same level are ?horizontal transformations?, and transformations between different levels are called ?vertical transformations?. There are 3 model transformations defined in $\pi$-SOD-M: $\pi$-UseCase to $\pi$--ServiceProcess; $\pi$--ServiceProcess to $\pi$-ServiceComposition; and $\pi$-ServiceComposition to $\pi$-PEWS.

%Figure \ref{fig:transformations} shows the transformation principle between the elements of the $\pi$-SCM meta-model used for representing the services' composition into the elements of the $\pi$-{\sc Pews} meta-model. There are two groups of rules: those that transform services' composition elements of the $\pi$-SCM to $\pi$-{\sc Pews} meta-models elements; and those that transform rules grouped by policies into {\em A-policy} types.

@@

The PIM to PIM transformation rules between $\pi$-UseCase and $\pi$-ServiceProcess meta-models (Table \ref{fig:transformationsUseCase-ServiceProcess})  specify how to express  simple and composite use cases defining the functions provided by the developed application, as a service process describing the application logic and its associated non-functional properties.  The principle is to express a set of $\pi$-use cases (i.e., use cases with constraints) related by associations of type {\sc Extend}   in terms of  actions related by a control flow and associated contracts.
\begin{figure}
\centering{
\includegraphics[width=0.96\textwidth]{figs/Table5}}
\caption{ $\pi$-UseCase to $\pi$-ServiceProcess transformation}
\label{fig:transformationsUseCase-ServiceProcess}
\end{figure}

Every {\sf Use Case} is transformed into an {\sf Action} defined in the $\pi$-ServiceProcess meta-model, and every {\sf Extend} association identified in a $\pi$-use case model is transformed into a  {\sf Fork node}.
If the extend association has only one {\sf use case},  the {\sf action} of the fork is an alternative flow. If the {\sf Extend} association has several source {\sf use cases},  different {\sf Actions} will mutual alternative flows of the fork node.  As shown in the $\pi$-ServiceProcess  meta-model an {\sc Activity service}  consists of a composition of one or more entities of type {\sc Action}.  {\sc Contracts}  are associated to  {\sc Actions}.

@@

   A {\sf Constraint} of a (composite) {\sf Use Case}  is transformed into an  {\sf Assertion}.  The set of constraints  in a $\pi$-UseCase model  transformed into assertions are grouped in a {\sf Contract} in a $\pi$-ServiceProcess model (figure  \ref{fig:transformationsUseCase-ServiceProcessRules}-a).
A constraint of a given type ({\sf Constraint type}) can have different transformations as stated in the following rules (figure   \ref{fig:transformationsUseCase-ServiceProcessRules}-b):
\begin{itemize}
\item  {\sc Business Constraint type} is transformed into any {\sf Assertion} that does not consider value attributes information like maxValue and minValue;
\item   {\sc Value Constraint type} is with the  {\sf  isExceptionalBehaviour}  set to false, is transformed into one {\sf Assertion};
\item  {\sc Value Constraint type}  with the {\sf  isExceptionalBehaviour } attribute set to true, is transformed into  {\sf Exceptional behaviour};
\end{itemize}
%Given the $\pi$-UseCase meta-model, to all {\sc Constraint} entity related with a {\sc Use Case}, there is a {\sc Contract} that compounds a set of {\sc Assertions} entity  and the {\sc Use Case} is refined in a service {\sc Action}.

 For {\sf Value Constraint Type}s
  the designer must specify value thresholds associated to variables according to the  { constraint}s  and {use case}s. By default, value constraints are transformed into pre-conditions and business constraints are transformed into post-conditions. This rule may be adjusted.

@@

The transformations for {\sc Extend} and {\sc Include} dependence elements are not as simple as the previous transformations (figures \ref{fig:transformaton-examples}-a). The workflow generated for each «extends» relationship with just two use cases (figure  \ref{fig:transformaton-examples}-a) is described by figure \ref{fig:transformaton-examples}-c). The generated workflow contains: one {\sc Fork node}, one {\sc Join node}, and four {\sc Control flow} elements, and also two {\sc Action} elements, one for each {\sc Use case}, if there are less than 2 extended {\sc Use case}. When there is more than one «extends» relationship among different use cases (figures \label{fig:transformaton-examples}-b and d), the transformation is proceed as: adding two more {\sc Control flow} for each new {\sc Use case}, and one {\sc Action} for each new extended {\sc Use case}.

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusicT3}
Considering the scenario example, it is possible to apply the same rule for the "publish music" use case, which has two extended use cases, "public twitter" and "public Facebook", applying the illustration expressed in figures \ref{fig:transformaton-examples}-b and d.
\end{example}

@@

For the {\sc Include} use case elements, the transformation is represented as a {\sc Action} sequence, as shown in figure \ref{fig:transformationsUseCase-ServiceProcessRules}-b. For each {\sc Use case} element an {\sc Action} element is generated. For a set of n {\sc Use case} we generate n-1 {\sc Object flow} elements. Each {\sc Control flow} links two {\sc Actions}.

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusicT4}
Given the "download music" use case from the scenario example, it includes pay- ment process to buy the music. It is represented as a include in the $\pi$-UseCase model and is transformed in a sequence flow in the $\pi$-ServiceProcess model (as presented in figure \ref{fig:transformation-example-include}). 
\end{example}

@@

% _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ .
\subsubsection{From $\pi$-ServiceProcess to $\pi$-ServiceComposition}
% _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ .
%The {\em A-policies} defined for the elements of the $\pi$-SCM are transformed into {\sc A-Policy} classes, named according to the names expressed in the source model. The transformation of the rules expressed in the $\pi$-SCM is guided by the event types associated to these rules.   The variables associated to an {\em A-policy} expressed in the $\pi$-SCM as {\sc\em $<$Variable:name, Variable:type$>$} are transformed into elements of type {\sc Variable} with attributes {\sc name} and {\sc type} directly specified from the elements {\sc\em  Variable:name} and {\sc\em Variable:type} of the $\pi$-SCM model.

@@

%Figure \ref{fig:pewsexpression} shows the correspondence between the model and the statements that implement it, with a schematic representation of the business process.
%\begin{figure}
%\centering{
%\includegraphics[width=0.85\textwidth]{figs/pews-expression}}
%\caption{Pews program implementing the "To Publish Music" application}
%\label{fig:pewsexpression}
%\end{figure}
Table \ref{fig:ServiceProcess-ServiceComposition}  describes the transformations between $\pi$-ServiceProcess and $\pi$-ServiceComposition meta-models.
Figure \ref{fig:ServiceProcess-ServiceComposition-Rules} presents the main transformation rules for this level. The  goal of this transformation  is to group all {\sf Contracts} and {\sf Actions} into {\sf Policies} and {\sf Service activities}, respectively.   The  criterion is that all {\sc Contracts} from the same {\sc Non-functional requirement} will be transformed into one {\sc Policy} type (e.g., safety, performance, and reliability {\sc Policy} types).

@@

{\sf Actions}  are transformed into an {\sf Action} in the target model, and every {\sc Service activity} is transformed into a {\sf Service activity} in the target model. 
For the non-functional specifications, a {\sf Contract} entity with the same { non-functional requirement} in the source model ($\pi$-ServiceProcess model) is transformed into a {\sf Policy} in the target model ($\pi$-ServiceComposition model). Each {\sf Assertion} is transformed into a {\sf Rule:Condition} attribute in the target model, but if the {\sf Assertion} has a value type, the name and the attributes in the source model will be transformed into a {\sc Variable} in the target model. The other {\sf Assertion property} values remains unchanged. The {\sf Assertion: Aproperty} attribute can have different transformations, depending of the following values:
\begin{itemize}
\item {\sf Post-Conditions} are transformed into{\sf  Post} in a $\pi$-ServiceComposition model; 
\item {\sf Preconditions} are transformed into {\sf Pre} in a $\pi$-ServiceComposition model;
\item {\sf Timerestrictions} are transformed into {\sf Time} in a $\pi$-ServiceComposition model.
\end{itemize}

@@

The {\sc Exceptional behaviour} entities are transformed into an {\sf Action} in a $\pi$- ServiceComposition model, and every {\sf Non-functional attribute} associated to an element ({\sf Contract} and {\sf Non-functional requirement}) in the $\pi$-ServiceProcess model becomes a {\sf Non-functional attribute} associated to the corresponding element ({\sf Policy}) in the $\pi$-ServiceComposition model.

Each {\sc Assertion} of a {\sc Contract} ( $\pi$-UseCase model) is transformed into a {\sc Rule} in the $\pi$-ServiceProcess model. The set of {\sc Contracts} generates a {\sc Policy} composed by those {\sc Rules}. Thus, the set of {\sc Rules} make up a single {\sc Policy}. It is important to highlight that only the {\sc Contracts} of the same {\sc Non-functional requirement} is transformed into the same {\sc Policy}. Different NFRs {\sc Contracts}, are transformed into different {\sc Policies}. Figure \ref{fig:ServiceProcess-ServiceComposition-Rules}-a show how this transformation is made.
Similar to the transformation between {\sc Contract} and {\sc Policy}, a set of {\sc Actions} is transformed into a single {\sc Service activity} entity. {\sc Actions} are related with {\sc Business collaborator}, and all information related to a {\sc Business Collaborator} are transformed from {\sc Packages} information ($\pi$-UseCase model).  {\sc Packages} are transformed into {\sc Business Collaborator}. Figure \ref{fig:ServiceProcess-ServiceComposition-Rules}-b presents the transformation between {\sc Action} and {\sc Service Activity} elements.
An exceptional behaviour is triggered by a failure in the verification of the condition of a rule. An {\sc Exceptional behaviour} is performed to preserve the state of the application. Figure \ref{fig:ServiceProcess-ServiceComposition-Rules}-c presents how an {\sc Exceptional behaviour} that can be transformed into an {\sc Action} call.

@@

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusicT5}
Considering the scenario example, the {\sc Action}?s update music {\sc Contract} is transformed is a {\sc Policy} with its {\sc Rules}. All contract {\sc Assertions} are transformed in RULE and its attributes, e.g. the login and password verification. The "securityLoginPolicy" is all set of {\sc Rules} that were transformed from the {\sc Assertions} in $\pi$-ServiceProcess model. The {\sc Non- functional requirement} information will be used to the {\sc Policy} generation comes from the initial use case model. Also the {\sc Business Collaborator} Facebook and Spotify information came from {\sc Package} $\pi$-UseCase entity element. All {\sc Contracts} of the same {\sc Non- functional requirement} are composed in a {\sc Policy}.
\end{example}

@@

Table \ref{fig:ServiceComposition-Pews-Rules} describes the PIM to PSM transformations between $\pi$-ServiceComposition and $\pi$-PEWS meta-models. We propose two groups of rules: those that transform services com- position elements of the $\pi$-ServiceComposition into $\pi$-PEWS meta-model elements; and those that transform rules grouped by policies into A-policy types.

A named action of the $\pi$-SCM represented by  {\sc\em Action} and {\sc\em Action:name} is transformed to a  named class {\sc Operation} with a corresponding attribute name {\sc Operation:name}. A  named service activity represented by the elements {\sc\em ServiceActivity}  and  {\sc\em ServiceActivity:name} of the $\pi$-SCM, are  transformed into a named operation of the $\pi$-{\sc Pews} represented by the elements  {\sc CompositeOperation} and {\sc CompositeOperation:name}. When more than one action is called, according to the following  composition patterns expressed using the operators {\sc\em merge, decision, fork and join} in the $\pi$-SCM the corresponding transformations, according to the PEWS operators presented above:
\begin{itemize}
\item   $op_1 . op_2$ if no {\sc\em ControlNode} is specified
\item ($op_1 \parallel op_2) . op_3$ if control nodes of type {\sc\em fork, join} are combined
 \item ($op_1 + op_2) . op_3$ if control nodes of type {\sc\em decision, merge} are combined
\end{itemize}

@@

The A-policies defined for the elements of the $\pi$-ServiceComposition are transformed into {\sc A-policy} classes, named according to the names expressed in the source model. The transformation of the rules expressed in the $\pi$-ServiceComposition is guided by the event types associated to these rules. The variables associated to an A-policy expressed in the $\pi$- ServiceComposition as $<$Variable:name, Variable:type$>$ are transformed into elements of type {\sc Variable} with attributes {\sf NAME} and {\sf TYPE} directly specified from the elements Variable:name and Variable:type of the $\pi$-ServiceComposition model.

As shown in Table \ref{fig:ServiceComposition-Pews-Rules}, for an event of type Pre the corresponding transformed rule is of type {\sc Precondition}; for an event of type Post the corresponding transformed rule is of type {\sc Post- condition}; finally, for an event of type TimeRestriction the corresponding transformed rule is of type {\sc Time}. The condition expression of a rule in the $\pi$-ServiceComposition (Rule:condition) is transformed into a class Condition:expression where the attributes of the expression are trans- formed into elements of type {\sc Attribute}.

Figure \ref{fig:Specific-Contract-Representation} presents the $\pi$-PEWS specification resulting from the $\pi$- ServiceComposition model transformation of our scenario example.

>>>>

@@

The PIM to PIM transformation rules specify how  to transform a $\pi$-use case model into  a $\pi$-service process model (Table \ref{fig:transformationsUseCase-ServiceProcess}). Given  (composite) $\pi$-use case model defining the functions of  the developed application and their associated constraints (representing business rules), the objective is to express them as a $\pi$-service process model describing the application logic and its associated non-functional properties.  The principle is to express a set of $\pi$-use cases (i.e., use cases with constraints) related by associations of type {\sc Extend}   in terms of  a business process consisting of actions related by a control flow and associated contracts.
\begin{figure}
\centering{
\includegraphics[width=0.96\textwidth]{figs/Table5}}
\caption{ $\pi$-UseCase to $\pi$-ServiceProcess transformation}
\label{fig:transformationsUseCase-ServiceProcess}
\end{figure}

As defined in the $\pi$-ServiceProcess  meta-model an {\sc Activity service}  consists of a composition of one or more entities of type {\sc Action}. 
Thus, every {\sf $\pi$-use case} is transformed into an {\sf Action} that will compose the target $\pi$-service process model.  Every {\sf Extend} relationship identified in a $\pi$-use case model is transformed into a  {\sf Fork node}.
If the extend relationship has: 
\begin{itemize}
\item only one {\sf use case} it is transformed into an {\sf action} of the fork in an alternative flow. 
\item several  {\sf use cases} are transformed as  different {\sf Actions} with mutual alternative flows in the fork node.   
%{\sc Contracts}  are associated to  {\sc Actions}.
\end{itemize}

@@

   A {\sf Constraint} of a (composite) {\sf Use Case}  is transformed into an  {\sf Assertion}.  The set of constraints  in a $\pi$-use case model  (transformed into assertions) are grouped in a {\sf Contract} in a $\pi$-ServiceProcess model (figure  \ref{fig:transformationsUseCase-ServiceProcessRules}-a).
A constraint of a given type ({\sf Constraint type}) can have different transformations as stated in the following rules (figure   \ref{fig:transformationsUseCase-ServiceProcessRules}-b). Given a  a {\sf Constraint} of type:

\begin{itemize}
\item   {\sc Business Constraint} is transformed into an {\sf Assertion} with no value attributes. %information like maxValue and minValue;
\item     {\sc Value Constraint}  with the  attribute {\sf  isExceptionalBehaviour}  set to false, is transformed into an {\sf Assertion};
\item  {\sc Value Constraint }  with the attribute {\sf  isExceptionalBehaviour }  set to true, is transformed into  {\sf Exceptional behaviour};
\end{itemize}
%Given the $\pi$-UseCase meta-model, to all {\sc Constraint} entity related with a {\sc Use Case}, there is a {\sc Contract} that compounds a set of {\sc Assertions} entity  and the {\sc Use Case} is refined in a service {\sc Action}.

 For transforming constraints of type {\sf Value Constraint }
  the designer must specify value thresholds associated to attributes as specified by  { constraint}s  and {use case}s. By default, value constraints are transformed into pre-conditions and business constraints are transformed into post-conditions.

@@

%The transformations for {\sc Extend} and {\sc Include} dependencies  are not as simple as the previous transformations (figures \ref{fig:transformaton-examples}-a).

The relationships of type  {\sc Extend} and {\sc Include}  determine the way the business process is expressed as a workflow.  The workflow generated for a relationship  «extends»  between  two use cases (figure  \ref{fig:transformaton-examples}-a) is described in figure \ref{fig:transformaton-examples}-c).  The generated workflow consists of  a {\sf Fork node}, a {\sf Join node},  four entities of type {\sc Control flow}, and also two entities of type {\sc Action} , one for each {\sf Use case}.

When there is more than one «extends» relationship among several use cases (figures \label{fig:transformaton-examples}-b and d), the transformation leads to  two more {\sc Control flow} entities for each  {\sc use case}, and an {\sf Action} for each  extended {\sf use case}.

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusicT3}
Considering the scenario example, it is possible to apply the same rule for the "publish music" use case, which has two extended use cases, "public twitter" and "public Facebook", applying the rules shown in figures \ref{fig:transformaton-examples}-b and d.
\end{example}

@@

 {\sf Include} use case entities, are transformed into an {\sf Action} sequence, as shown in figure \ref{fig:transformationsUseCase-ServiceProcessRules}-b. A {\sc Use case} element is transformed into an {\sf Action}. For a set of n {\sc Use case} is transformed into an  n-1 {\sf Object flow} elements. Each {\sf Control flow} is transformed into two {\sf Actions} associated by a relationship.

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusicT4}
Given the "download music" use case of the scenario example, it includes the payment process to buy the music. It is represented as a include in the $\pi$-UseCase model and it is transformed in a sequence flow in the $\pi$-ServiceProcess model (as shown in figure \ref{fig:transformation-example-include}). 
\end{example}

@@

% _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ .
\subsection{From $\pi$-ServiceProcess to $\pi$-ServiceComposition}
% _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ . _ .
%The {\em A-policies} defined for the elements of the $\pi$-SCM are transformed into {\sc A-Policy} classes, named according to the names expressed in the source model. The transformation of the rules expressed in the $\pi$-SCM is guided by the event types associated to these rules.   The variables associated to an {\em A-policy} expressed in the $\pi$-SCM as {\sc\em $<$Variable:name, Variable:type$>$} are transformed into elements of type {\sc Variable} with attributes {\sc name} and {\sc type} directly specified from the elements {\sc\em  Variable:name} and {\sc\em Variable:type} of the $\pi$-SCM model.

@@

%Figure \ref{fig:pewsexpression} shows the correspondence between the model and the statements that implement it, with a schematic representation of the business process.
%\begin{figure}
%\centering{
%\includegraphics[width=0.85\textwidth]{figs/pews-expression}}
%\caption{Pews program implementing the "To Publish Music" application}
%\label{fig:pewsexpression}
%\end{figure}
Table \ref{fig:ServiceProcess-ServiceComposition}  shows the transformation rules for transforming a $\pi$-Service process model into a   $\pi$-ServiceComposition model.
As illustrated in Figure \ref{fig:ServiceProcess-ServiceComposition-Rules} the  principle of this transformation  is to group respectively  {\sf Contracts}  and {\sf Actions} into {\sf Policies} and {\sf Service activities}.

Each {\sf Assertion} of a {\sf Contract} in a $\pi$-UseCase model is transformed into a {\sf Rule} in a $\pi$-ServiceProcess model. The set of {\sf Contracts} generates a {\sf Policy} composed by those {\sf Rules}. Thus, the set of {\sf Rules} make  a single {\sf Policy}. Only the {\sc Contracts} of the same {\sc Non-functional requirement} are transformed into the same {\sc Policy}. Different NFRs {\sc Contracts}, are transformed into different {\sc Policies}. Figure \ref{fig:ServiceProcess-ServiceComposition-Rules}-a show how this transformation is made (e.g., safety, performance, and reliability {\sc Policy} types).

@@

{\sf Actions}  of a $\pi$-service process model are transformed into an {\sf Action} in a $\pi$-service composition model, and every {\sf Service activity} of a $\pi$-service process model is transformed into a {\sf Service activity} in a $\pi$-service composition model (see Figure \ref{fig:ServiceProcess-ServiceComposition-Rules}-b). {\sf Actions} are related to a  {\sf Business collaborator}. A {\sf Package}  in the $\pi$-use case model is transformed   into a {\sf Business Collaborator}.  
 {\sf Contract}s  concerned with the same { non-functional requirement} in a $\pi$-service process model are transformed into a {\sf Policy} in a $\pi$-service composition model. Each {\sf Assertion} of a {\sf Contract} is transformed into a {\sf Rule:Condition} attribute. If the {\sf Assertion} has a value type, the name and the attributes are transformed into a {\sf Variable} in the target model.  The {\sf Assertion: Aproperty} attribute can have different transformations, according to  the following rules:
\begin{itemize}
\item a {\sf Post-Condition} is transformed into a {\sf  Post}. 
\item a {\sf Precondition} is transformed into {\sf Pre}.
\item a {\sf Timerestriction} is transformed into {\sf Time}.
\end{itemize}

@@

An {\sf Exceptional behaviour} entity is transformed into an {\sf Action} in a $\pi$-service composition model (see Figure \ref{fig:ServiceProcess-ServiceComposition-Rules}-c ), and every {\sf Non-functional attribute} associated to an element ({\sf Contract} and {\sf Non-functional requirement}) in the $\pi$-ServiceProcess model becomes a {\sf Non-functional attribute} associated to the corresponding element {\sf Policy} in the $\pi$-ServiceComposition model.

@@

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusicT5}
Considering the scenario example, 
%the {\sf Action}s update music {\sc Contract} is transformed is a {\sc Policy} with its {\sc Rules}. All contract {\sc Assertions} are transformed in RULE and its attributes, e.g. the login and password verification. 
the "securityLoginPolicy" consists of a set of {\sf Rules} that were transformed from the {\sf Assertions} in $\pi$-ServiceProcess model. 
%The {\sc Non- functional requirement} information will be used to the {\sc Policy} generation comes from the initial use case model. Also 
The {\sf Business Collaborator} Facebook and Spotify information came from entity of type {\sc Package}  in the $\pi$-use case model. 
%All {\sc Contracts} of the same {\sc Non- functional requirement} are composed in a {\sc Policy}.
\end{example}

@@

Table \ref{fig:ServiceComposition-Pews-Rules} describes the PIM to PSM transformations between a $\pi$-Service composition and a $\pi$-PEWS program. We propose two groups of rules: those that transform services composition entities of a $\pi$-Service composition model into $\pi$-PEWS model ; and those that transform rules grouped by policies into A-policy types.

A named action of the $\pi$-SCM represented by  {\sf Action} and {\sf Action:name} is transformed to an  {\sf Operation} with a corresponding attribute name {\sf Operation:name}. A  named service activity represented by the elements {\sf ServiceActivity}  and  {\sf ServiceActivity:name} of the $\pi$-SCM, are  transformed into a named operation of the $\pi$-{Pews} represented by the entities  {\sf CompositeOperation} and {\sf CompositeOperation:name}. When more than one action is called, according to the following  composition patterns expressed using the operators {\sc\em merge, decision, fork and join} in the $\pi$-SCM the corresponding transformations, according to the PEWS operators presented above:
\begin{itemize}
\item   $op_1 . op_2$ if no {\sc\em ControlNode} is specified
\item ($op_1 \parallel op_2) . op_3$ if control nodes of type {\sc\em fork, join} are combined
 \item ($op_1 + op_2) . op_3$ if control nodes of type {\sc\em decision, merge} are combined
\end{itemize}

@@

The A-policies defined for the entities of a $\pi$-service composition model are transformed into {\sf A-policy} entities, named according to the names expressed in the source model. The transformation of the rules expressed in a $\pi$-service composition is guided by the event types associated to these rules. The variables associated to an A-policy expressed in a $\pi$-service composition model as {\sf $<$Variable:name, Variable:type$>$} are transformed into entities  {\sf Variable} with attributes {\sf Name} and {\sf Type} directly specified from the elements {\sf Variable:name} and {\sf Variable:type} of a $\pi$-service composition model.

As shown in Table \ref{fig:ServiceComposition-Pews-Rules}, for an event of type Pre the corresponding transformed rule is a {\sf Precondition}; for an event of type Post the corresponding transformed rule is a {\sf Post- condition}; finally, for an event of type TimeRestriction the corresponding transformed rule is a {\sf Time}. The condition expression of a rule in a $\pi$-service composition ({\sf Rule:condition}) is transformed into a  {\sf Condition:expression} where the attributes of the expression are transformed into  entities {\sf Attribute}.

Figure \ref{fig:Specific-Contract-Representation} shows the $\pi$-PEWS code resulting from the $\pi$- service composition model  of our scenario example.
